\section{Fingerprinting}
A string fingerprint can be understood as the answer vector to a small
collection of yes/no questions asked of a string. In the most precise
form one could imagine, there would be one such question per
\emph{atomic feature} of interest---for instance, ``does this string
contain the letter \texttt{a}?'', ``does it contain the bigram
\texttt{th}?'', ``does it start with a digit?'', or any other Boolean
property one cares to evaluate. Evaluating every atomic feature gives a
lossless, but potentially very large, summary of the string.

The fingerprint trades this precision for size. Instead of storing one
bit per atomic feature, we partition the features into $n$ groups and
\emph{or}~them together: a single bit fires whenever \emph{any} of its
constituent features fires. Each bit of the fingerprint is therefore
not a single feature but the disjunction of many features. This
compression is what makes fingerprints small and fast to evaluate---and
also what makes them lossy. The design problem is then twofold: choose
the family of atomic features to draw from, and choose how to group
them into bits.

Let $\mathcal{F} = \{f_1, f_2, \dots, f_m\}$ be a finite family of
\emph{atomic features}, where each $f_i$ is a Boolean function on
strings:
\begin{equation*}
  f_i : \Sigma^{*} \to \{0,1\}.
\end{equation*}
We make no assumption about what each $f_i$ computes; it may test for
the presence of a character, the presence of an $n$-gram, a positional
condition, the output of a hash function, or any other property.

A fingerprint of bitwidth $n$ is induced by a partition
$\{B_1, \dots, B_n\}$ of $\mathcal{F}$. The $j$-th bit of the
fingerprint of a string $s$ is the disjunction of the features assigned
to bin $B_j$:
\begin{equation}
  \Vector{f}_s[j] \;=\; \bigvee_{f \in B_j} f(s).
  \label{eq:bit-as-or}
\end{equation}
The full fingerprint $\Vector{f}_s \in \{0,1\}^n$ is the vector of these
disjunctions.

\paragraph{Example.} Let $\Sigma = \{a, e, h, l, n, t, u\}$ and take
the family $\mathcal{F} = \{f_a, f_e, f_h, f_l, f_n, f_t, f_u,
g_{\text{th}}, g_{\text{ut}}\}$, where $f_c(s) = \mathbb{1}[c \in s]$
and $g_{xy}(s) = \mathbb{1}[xy \text{ occurs in } s]$. With bitwidth
$n = 3$ and partition $B_1 = \{f_a, f_l, f_u\}$, $B_2 = \{f_e, f_h,
f_n, f_t\}$, $B_3 = \{g_{\text{th}}, g_{\text{ut}}\}$, the
fingerprints of a few strings are:
\begin{center}
  \begin{tabular}{lccc}
    \toprule
    String & $\Vector{f}[1]$ & $\Vector{f}[2]$ & $\Vector{f}[3]$ \\
    \midrule
    \texttt{utn}   & 1 & 1 & 1 \\
    \texttt{lute}  & 1 & 1 & 1 \\
    \texttt{heat}  & 1 & 1 & 0 \\
    \texttt{ten}   & 0 & 1 & 0 \\
    \bottomrule
  \end{tabular}
\end{center}
For instance, \texttt{lute} sets $\Vector{f}[3] = 1$ via the bigram \texttt{ut};
the fingerprint records that \emph{some} feature in $B_3$ fired, but
not which.


\subsection{Approximate Filtering}

Fingerprints become useful when we want to quickly decide whether a
candidate string $s$ could possibly contain a pattern $\zeta$, without
performing the full containment test.

\paragraph{The pruning rule.} For a pattern $\zeta$ and a candidate string
$s$, the fingerprint check accepts $s$ iff
\begin{equation}
  \Vector{f}_\zeta \subseteq \Vector{f}_s
  \;\;\Longleftrightarrow\;\;
  \forall j \in [n]:\; \Vector{f}_\zeta[j] = 1 \Rightarrow \Vector{f}_s[j] = 1.
  \label{eq:subset-check}
\end{equation}
For this check to be a sound filter---never rejecting a true
match---the feature family $\mathcal{F}$ must satisfy a
\emph{monotonicity} condition: whenever $\zeta$ is contained in $s$ in the
sense the query intends, every feature satisfied by $\zeta$ must also be
satisfied by $s$,
\begin{equation}
  \zeta \sqsubseteq s \;\;\Longrightarrow\;\; \forall f \in \mathcal{F}:\;
  f(\zeta) \le f(s).
  \label{eq:monotonicity}
\end{equation}
Under~\eqref{eq:monotonicity}, the subset
check~\eqref{eq:subset-check} cannot produce false negatives.

\paragraph{Where false positives come from.} The converse of the
subset check fails because of the \emph{or}-compression
in~\eqref{eq:bit-as-or}. A string $s$ may light bit $j$ via a feature
$f \in B_j$ unrelated to those the pattern requires; the fingerprint
cannot tell which feature of $B_j$ caused the bit to fire. Two design
choices govern how often this happens:
\begin{enumerate}
  \item The \textbf{feature family} $\mathcal{F}$. Richer families
  (e.g., $n$-grams, positional features) carry more information per
  atomic feature and therefore tolerate coarser bin assignments, at
  the cost of a larger $|\mathcal{F}|$ and a more expensive
  partition-selection problem.
  \item The \textbf{partition} $\{B_1, \dots, B_n\}$. Grouping
  features the workload needs to \emph{distinguish} into the same bin
  destroys that information; grouping interchangeable features wastes
  little.
\end{enumerate}

\paragraph{Example.} Reusing the setup from above, take pattern $\zeta =
\texttt{utn}$ with $\Vector{f}_\zeta = (1, 1, 1)$. Then \texttt{lute} passes the
filter as a false positive (its $\Vector{f}[3]$ bit is lit by \texttt{ut} rather
than by \texttt{utn}), while \texttt{heat} is correctly rejected on
$\Vector{f}[3]$ and \texttt{ten} on $\Vector{f}[1]$. Replacing $B_3$ with another
character bin would weaken the filter on \texttt{heat}; merging
\texttt{u} and \texttt{t} into one bin would weaken it on
\texttt{ten}.


Now building on this tool, we design a PIOP where 

\begin{DefinePIOP}{Approximate Filtering Check}{approx-filer}
    \begin{enumerate}[nosep]
        \item $\PIOPProver$ and $\PIOPVerifier$ compute the fingerprint of the pattern $\zeta$ as $\Vector{f}_\zeta$ and the field element encoding of the fingerprint as $f_\zeta=\sum_{j=0}^{n-1} 2^j \Vector{f}_\zeta[j]$.
        \item $\PIOPProver$ divides the fingerprint col
    \end{enumerate}
\end{DefinePIOP}

\subsection{Exact Filtering}


\subsubsection{Activator Consistency Check}
The \emph{Activator Consistency Check} takes the source column $\mathsf{src}$, the data activator $\mathsf{a_d}$, the length column $\mathsf{l}$, and the string activator $\mathsf{a_h}$, and proves that $\mathsf{a_d}$ and $\mathsf{a_h}$ are consistent: for every string, the number of its active characters equals its length when the string is active, and is $0$ otherwise. Equivalently, for every string index $i$,
\begin{displaymath}
  \mathsf{a_h}[i]\cdot\mathsf{l}[i] \;=\; \sum_{j:\,\mathsf{src}[j]=i} \mathsf{a_d}[j].
\end{displaymath}
We package this as the relation
\begin{displaymath}
\NPRelation_{\mathsf{ActCons}} = \left\{\,
  \NPInstance = \bot,\;
  \NPInstancey = (\Oracle{\MLE{\mathsf{src}}}, \Oracle{\MLE{\mathsf{a_d}}}, \Oracle{\MLE{\mathsf{l}}}, \Oracle{\MLE{\mathsf{a_h}}}),\;
  \NPWitness = (\mathsf{src}, \mathsf{a_d}, \mathsf{l}, \mathsf{a_h})
  \;\middle|\;
  \forall i,\; \mathsf{a_h}[i]\cdot\mathsf{l}[i] = \!\!\!\sum_{j:\,\mathsf{src}[j]=i}\!\!\! \mathsf{a_d}[j]
\,\right\}.
\end{displaymath}
The check is exactly a group-by-sum: grouping the rows $(\mathsf{src}, \mathsf{a_d})$ by $\mathsf{src}$ and summing $\mathsf{a_d}$ within each group must reproduce $\mathsf{l}$ on the active strings (those marked by $\mathsf{a_h}$). Treating $(\mathsf{src}, \mathsf{a_d})$ as an input table grouped by its key and $(\mathsf{ind}, \mathsf{l})$ with activator $\mathsf{a_h}$ as the grouped output, this is discharged by a single \cref{piop:keyedsumcheck}.
\begin{DefinePIOP}{Activator Consistency Check}{act-cons}
  \emph{Inputs:} $\Oracle{\MLE{\mathsf{src}}}, \Oracle{\MLE{\mathsf{a_d}}}, \Oracle{\MLE{\mathsf{l}}}, \Oracle{\MLE{\mathsf{a_h}}}$.
  \begin{enumerate}[nosep]
    \item $\PIOPProver$ and $\PIOPVerifier$ locally form the input table $\Left$ with key column $\Left.\key = \mathsf{src}$, value column $\Left.\val = \mathbf{1}$, and activator $\Left.\act = \mathsf{a_d}$ (so the keyed sum counts the active characters of each string).
    \item $\PIOPProver$ and $\PIOPVerifier$ locally form the output table $\Right$ with key column $\Right.\key = \mathsf{ind}$ (the index column, $\mathsf{ind}[i] = i$), value column $\Right.\val = \mathsf{l}$, and activator $\Right.\act = \mathsf{a_h}$.
    \item $\PIOPProver$ and $\PIOPVerifier$ invoke \cref{piop:keyedsumcheck} on $\Left$ and $\Right$, which forces $\sum_{j:\,\mathsf{src}[j]=i}\mathsf{a_d}[j] = \mathsf{a_h}[i]\cdot\mathsf{l}[i]$ for every key $i$.
  \end{enumerate}
\end{DefinePIOP}



\begin{table}[h!]
\centering
\small
\begin{tabular}{p{3.8cm}p{4.5cm}p{5cm}}
\toprule
\textbf{SQL String Operator} & \textbf{Approx. Filtering Applicable?} & \textbf{Used in TPC-H?} \\
\midrule
\texttt{LIKE '\%pat\%'} (substring)
  & Yes --- canonical case; character/$n$-gram features are monotone
  & Yes: Q9, Q13 (negated), Q14, Q16 (negated) \\
\addlinespace
\texttt{LIKE 'pat\%'} (prefix)
  & Yes --- with anchored/positional features (or sortedness)
  & Yes: Q2, Q20 \\
\addlinespace
\texttt{LIKE '\%pat'} (suffix)
  & Yes --- with end-anchored features
  & No \\
\addlinespace
\texttt{LIKE 'pat'} (exact, no wildcards)
  & Trivially yes, but equality is usually faster directly
  & No (TPC-H uses \texttt{=} instead) \\
\addlinespace
\texttt{LIKE 'p\_t'} (single-char wildcard)
  & Partial --- anchor what you can; treat \texttt{\_} as unknown
  & No \\
\addlinespace
\texttt{LIKE '\%a\%b\%'} (multi-fragment)
  & Yes --- filter on each required fragment, AND the results
  & Yes: Q13 \\
\addlinespace
\texttt{NOT LIKE}
  & No directly --- monotonicity runs the wrong way; filter the positive form and complement
  & Yes: Q13, Q16 \\
\addlinespace
\texttt{=} (exact equality)
  & Trivially yes, but direct comparison or hashing wins
  & Pervasive \\
\addlinespace
\texttt{<}, \texttt{>}, \texttt{<=}, \texttt{>=} (lex.)
  & Partial --- works with order-preserving features; not the technique's strength
  & No (used on numeric/date columns only) \\
\addlinespace
\texttt{IN (lit1, lit2, ...)} (string set)
  & Yes --- disjunction over members; equivalent to multiple equality filters
  & Yes: Q19, Q22 \\
\addlinespace
\texttt{NOT IN} (string set)
  & No directly --- same issue as \texttt{NOT LIKE}
  & Yes: Q16 \\
\addlinespace
\texttt{BETWEEN} (lex.)
  & Partial --- same caveats as \texttt{<}/\texttt{>} on strings
  & No (numeric/date use in TPC-H) \\
\addlinespace
\texttt{SUBSTRING} / \texttt{SUBSTR} extract
  & N/A as a predicate; filter the enclosing \texttt{=} or \texttt{LIKE}
  & Yes: Q22 (\texttt{SUBSTRING(c\_phone FROM 1 FOR 2)}) \\
\addlinespace
\texttt{CONCAT} / \texttt{||}
  & N/A --- not a predicate
  & No (in \texttt{WHERE} clauses) \\
\addlinespace
Regex (\texttt{SIMILAR TO}, \texttt{\textasciitilde})
  & Partial --- works when required literal fragments can be extracted
  & No \\
\bottomrule
\end{tabular}
\caption{Applicability of approximate filtering to SQL string operators, and their occurrence in TPC-H queries.}
\label{tab:sql-ops-filter}
\end{table}